---
title: "Lab 5"
author: "Anisha Babu"
date: "October 21, 2020"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE) 
```

1. Create a RStudio Project on your machine.

* Within the Project create a folder for *data* and *scripts*.
* Download the following three datasets into the project *data* folder.
  + *sesame13.sav*
  + *star.csv*
  + *ais.xlsx*

2. Read in the following three datasets, using `{here}` and `{rio}`:
  * *sesame13.sav*
  * *star.csv*
  * *ais.xlsx*
  
**Hint**: For the *ais.xlsx* dataset look at the `skip` argument within the `rio` help documentation.
  
```{r import}
library(tidyverse)
library(dplyr)
library(here)
library(rio)
ais <- import(here("data", "ais.xlsx"), skip = 15)
sesame13 <- import(here("data", "sesame13.sav"))
star <- import(here("data", "star.csv"))

```

3. Using the `ais` data, compute the average red blood cell count and average bmi by sport. Output these data as SPSS and EXCEL files. 

```{r q2}
by_sport <- group_by(ais,sport)
avg_rbc <- summarize(by_sport, avg_rcc = mean(by_sport$rcc))
avg_bmi <- summarize(by_sport, avg_bmi = mean(by_sport$bmi))
export(avg_rbc, here("data", "avg_rbc.sav"))
export(avg_bmi, here("data", "avg_bmi.sav"))
export(avg_rbc, here("data", "avg_rbc.xlsx"))
export(avg_bmi, here("data", "avg_bmi.xlsx"))



```

4. Use the `sesame` dataset to answer the following question: Was the average female age higher in schools or at home? 

```{r q3}
female_school <- filter(sesame13, sex == 2, setting == 1) 
avg_school <- mean(female_school$age)
female_home <- filter(sesame13, sex == 2, setting == 0) 
avg_home <- mean(female_home$age)
#Average age was higher in schools


```

5. First, how many rows and columns are in the *star* dataset? Next, remove outliers using a really poor method, just for practice, by eliminating students whose math (*tmathss*) scores were more than three standard deviations **above or below** the corresponding mean. How many rows are in the dataset now?

```{r q4}
dim(star)
#star has 5748 rows and 12 columns

avg_math <- mean(star$tmathss)
sd3_math <- 3 * (sd(star$tmathss))
min_math <- avg_math -sd3_math
max_math <- avg_math + sd3_math

star <- star[!star$tmathss > max_math & !star$tmathss < min_math,]

dim(star)
#there are now 5743 rows 

  
```

6. Use the `star` dataset to compute standardized math and reading scores; name these variables *tmathss* and *treadss*. This means, for each of these variables, subtract the mean from each observation and divide by the standard deviation: $x_s = \frac{x_i - \bar{X}}{sd(X)}$. 
  + Check that your computations were correct by computing the mean and standard deviation of each variable (they should be 0 and 1). 
  + Compute the mean of the standardized variable for all *sex*/*frl* combinations (I'm asking you to extend what you know here, we haven't talked explicitly about how to do this yet, but you have seen examples). 
  + What do you make of these findings? Do you see an effect by *sex*? An *FRL* effect? Is there evidence of an interaction (i.e., that the effect of *FRL* is greater for boys versus girls)?

```{r q5}
avg_math <- mean(star$tmathss)
sd_math <- sd(star$tmathss)
avg_read <- mean(star$treadss)
sd_read <- sd(star$treadss)
star$tmathss <- (star$tmathss - avg_math) / sd_math
star$treadss <- (star$treadss - avg_read) / sd_read

avg_math2 <- mean(star$tmathss)
sd_math2 <- sd(star$tmathss)
avg_read2 <- mean(star$treadss)
sd_read2 <- sd(star$treadss)

star %>%
  group_by(sex, frl) %>%
  summarize(avg_math = mean(tmathss), avg_read = mean(treadss))

#this summary does not appear to show an effect of sex 
#frl is associated with negative scores
#the magnitude of score is due to interaction (higher magnitude of boys with frl and girls without frl)

```

Please submit you .Rmd file, as well as the rendered .html file.